A Study of Residual Adapters for Multi-Domain Neural Machine Translation

### 2020-10-09 2020/wmt_20/mdadapt.tex
Fin des revisions

We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training. Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyTodo{Answer question of reviewer 3 about splits ?}
   %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.
\fyFuture{Is this part on significance testing still accurate ?}

@@

For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells. Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$. \revision{\citet{Bapna19simple} showed that the performance of adapted models increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer. We used the same setting in our experiments.}\fyDone{Check: is it $b$?}

% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers. \revision{For the \system{Mixed} model, we use an initial learning rate of 1.0 and take the concatenation of the validation sets of 6 domains for development. In the fine-tuning experiments, we continue training using \system{Mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps. In the multi-task training, we use the same learning rate schedule as for \system{Mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.}\fyDone{Describe the block adaptation layer - voir slides}

@@

\begin{table*}[t!]
  \centering
  \begin{tabular}{|p{3.5cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}        & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \hline 
    \system{DC-Tag}      & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}     & 37.7 & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9  \\
    \system{LDR}            & 37.0  & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8    \\
    \system{TTM}           & 37.3  & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
    \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
    \hline
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
    \system{Res-Adap-MT}$^{+}$ & 37.5 & 57.1 & 52.4 & 33.7 & 46.2 & 89.5 & 52.7 \\
    \hfill {Res-Adap-MT} (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 \\
    \hline
  \end{tabular}
  \caption{Translation performance of various multi-domain MT systems (En$\rightarrow$Fr) compared to variants of the residual adapter models.}
  \label{tab:performance-multi}
\end{table*}
The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two-step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch. We also report results for the same system, using the the parameters of the \system{Mixed} model as initialization (\system{Res-Adap-MT}${^+}$).

@@

% Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}. However, \system{MDL-Res} is outperformed by \system{Res-Adap}. It means that independently fine tuning the generic models is better than multi-task training. However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
\revision{Our comparison of training regimes (\system{Res-Adap}  vs   \system{Res-Adap-MT} vs \system{Res-Adap-MT}$^+$) suggests that the simultaneous learning of all parameters is detrimental to performance in our settings, where the 2-step procedure implemented in \system{Res-Adap}) yield the best scores, even when initializing with good parameter values . This may be because in this setting, the adapters have access to a stable version of the generic system. The last line ({Res-Adap-MT} (gen)) gives the results for \system{Res-Adap-MT} trained system in which we cancel the adaptor in inference - comparing this to \system{Mixed} show how differently the generic parts of these two systems behave.}

>>>>

We randomly select in each corpus a development and a test set of 1,000 lines each and keep the rest for training.\footnote{Scripts to replicate these experiments are available at url{}.}\fyTodo{github} Development sets help choose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings}\fyDone{Answer question of reviewer 3 about splits ?}
   %    Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.
\fyFuture{Is this part on significance testing still accurate ?}

@@

For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells. Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$. \citet{Bapna19simple} showed that the performance of adapted models increases with respect to the size of the inner dimension and obtained performance close to the full fine-tuned model with $b=1024$, which is twice as large as the dimension of a Transformer layer. We used the same setting in our experiments.\fyDone{Check: is it $b$?}

% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers. For the \system{Mixed} model, we use an initial learning rate of $1.0$ and take the concatenation of the validation sets of 6~domains for development. In the fine-tuning experiments, we continue training using \system{Mixed} as starting point, using the same learning rate schedule, and continuing the incrementation of the number of steps. In the multi-task training, we use the same learning rate schedule as for \system{Mixed}: for each iteration, we sample a domain a probability proportional to its size; we then sample a batch of~12,288 tokens that is used to update the shared parameters and the parameters of the corresponding adapter.\fyDone{Describe the block adaptation layer - voir slides}

@@

\begin{table*}[t!]
  \centering
  \begin{tabular}{|p{3.5cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}        & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \hline 
    \system{DC-Tag}      & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}     & 37.7 & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9  \\
    \system{LDR}            & 37.0  & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8    \\
    \system{TTM}           & 37.3  & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
    \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
    \hline
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
    \system{Res-Adap-MT}$^{+}$ & 37.5 & 57.1 & 52.4 & 33.7 & 46.2 & 89.5 & 52.7 \\
    \hfill {Res-Adap-MT} (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 \\
    \hline
  \end{tabular}
  \caption{Translation performance of various multi-domain MT systems (En$\rightarrow$Fr) compared to variants of the residual adapter models.}
  \label{tab:performance-multi}
\end{table*}
The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two-step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch. We also report results for the same system, using the the parameters of the \system{Mixed} model as initialization (\system{Res-Adap-MT}${^+}$).\footnote{This system also includes a layer dropout policy that cancels adapter layers with probability $0.5$}

@@

% Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}. However, \system{MDL-Res} is outperformed by \system{Res-Adap}. It means that independently fine tuning the generic models is better than multi-task training. However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.
Comparing training schemes (\system{Res-Adap} vs \system{Res-Adap-MT} vs \system{Res-Adap-MT}$^+$) suggests that the simultaneous learning of all parameters is detrimental to performance in our settings: we see that the 2-step procedure implemented in \system{Res-Adap} always yields the best scores, even when \system{Res-Adap-MT} is initialized with good parameter values . This may be because in this setting, the adapters have access to a stable version of the generic system. The last line ({Res-Adap-MT} (gen)) gives the results for a \system{Res-Adap-MT} trained system in which we cancel the adapter in inference - comparing this to \system{Mixed} shows how differently the generic parts of these two systems behave.
